---
title: "Data Visualization with ggplot"
output: github_document
---

loading tidyverse
```{r}
library(tidyverse)
```

Creates an empty graph

```{r}
ggplot(data = mpg)
```

geom_point() adds a layer of points to your plot
aes() specify variables to map to x and y-axes

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

number of rows and columns

```{r}
nrow(mtcars)
ncol(mtcars)
```

documentation

```{r}
?mpg
```

a scatterplot of hwy versus cyl

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cyl, y = hwy))
```

scatterplot of class versus drv

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```

Scaling: Assigning unique level of aesthetic to each unique value of the variable.

Aesthetic color

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

Aesthetic size

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

Aesthetic aplha (for transparency)

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

Aesthetic shape

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

Manually setting aesthetic properties

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "purple")
```

Mapping a continuous variable to color and size.
Continuous variables cannot be mapped to shape

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = year))
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = year))
```
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = hwy, size = displ))
```

Mapping multiple aesthetics to same variable

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class, size = class, shape = class))
```

Stroke aesthetic

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), shape = 21, color = "purple", fill = "yellow", size = 4, stroke = 2)
```

Conditional operation

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```

Facet_wrap():
variable passed should be discrete

will only produce plots for the combinations of variables that have values

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```

facet_grid() will produce a grid of plots for each combination of variables that you specify, even if some plots are empty.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```

Facet on a continuous variable
plot contains a facet for each distinct value

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ cty)
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(~ cty)
```
drv ~ . facet by values of drv on the y-axis.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

. ~ cyl will facet by values of cyl on the x-axis

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

Geometrical Object
A geom is the geometrical object that a plot uses to represent data.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

geom_smooth()

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

linetype in geom_smooth()

geom_smooth() separates the cars into three lines based on their drv value

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

group aesthetic

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
```
```{r}
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```

Add multiple geom functions

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

More efficient way

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth()
```

filter

se: Computes standard error of the mean.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth(
    data = filter(mpg, class == 'subcompact'),
    se = FALSE
  )
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

Recreating graphs from exercise

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(size = 3) +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(size = 3) +
  geom_smooth(mapping = aes(group = drv), se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point(size = 3) +
  geom_smooth(se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv), size = 3) +
  geom_smooth(se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv), size = 3) +
  geom_smooth(mapping = aes(linetype = drv), se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(size = 3, color = 'white') +
  geom_point(mapping = aes(color = drv))
```

The algorithm used to calculate new values for a graph is called a stat.
default value for stat is “count”

geom_bar()

```{r}
ggplot(dat = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

This works because every geom has a default stat, and every stat has a default geom. 

```{r}
ggplot(dat = diamonds) +
  stat_count(mapping = aes(x = cut))
```

identity

```{r}
demo <- tribble(
  ~a,      ~b,
  "A", 20,
  "B", 35,
  "C", 15
)

ggplot(dat = demo) +
  geom_bar(mapping = aes(x = a, y = b), stat = 'identity')

```

Bar chart of proportion

```{r}
ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = cut, y = ..prop.., group = 1)
  )
```

stat_summary()

```{r}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```
```{r}
?stat_summary
```

```{r}
ggplot(data = diamonds) +
  geom_pointrange(
    mapping = aes(x = cut, y = depth),
    stat = 'summary',
    fun.min = min,
    fun.max = max,
    fun = median
  )
```

If group = 1 is not included, all the bars will have the same height.

The function geom_bar() assumes that the groups are equal to the x values since the stat computes the counts within the group.

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = ..prop..))
```

```{r}
ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = color, y = ..count.. / sum(..count..))
  )
```

